### 2018-06-11 express2018/part2.tex
Cumulative updates (still not touching the 2nd part)
Clean up, re-ordered part2 before adding new contents

In the second half of this paper, we present a formalization of CCS
including the new concepts and theorems proposed in the first half of
this paper. The main purpose is to convince the readers that, there's no flaw\footnote{Perheps this is the first
  time a frontier development in Concurrency Theory ships with its
  formalization in the same paper.} in the informal proofs. The
formalization work is based on a formalization of CCS in HOL theorem
prover (HOL4), described in second author's MSc thesis
\cite{Chun:2017} and now lives in the official examples folder of HOL4 source
code\footnote{\url{https://github.com/HOL-Theorem-Prover/HOL}}.
The work is essentially a porting and extension of the early CCS
formalization by Monica Nesi \cite{Nesi} in HOL88.

We will not repeat all definitions and theorems of CCS again by their
formalized versions. Instead, we just focus on several highlights in
this work, i.e.
\begin{enumerate}
\item The use of recursion operators in formalizing process constants
  and variables;
\item The formalization of semantics context in $\lambda$-expressions and the theory of
  (pre)congruence for CCS;
\item The definition and uses of trace in the proof of unique solution of
  contractions theorem;
\item The proof of hard version of ``coarsest congruence
  contained in $\approx$'' theorem.
\end{enumerate}

@@

\subsubsection{Limitations}

There're two major limitations in this work. One is the class of CCS
processes, i.e. we're limited with Finitary CCS which has no infinite
sums or parallel compositions. The only important theorem in which
arbitrary infinite sums of CCS processes must be used, is the following
``coarsest congruence contained in weak bisimilarity'' theorem without
any limitations on the involved processes:
\begin{equation}
\forall p\; q.\; p \approx^c\! q \Longleftrightarrow \forall r.\; p+r \approx
q+r
\end{equation}
Milner's classical proof of above result needs to assume that $p$ and $q$ didn't
use up all available labels. The newer proof by van Glabbek requires a
special choice of $r$ which is not weak bisimilar with any process
transited from $p$ and $q$, and the actual chosen process may contain
arbitrary large sums of CCS processes. This goes beyond the
expressiveness of HOL, as result we followed the same proof ideas but
limited with finite state CCS only.

The other limitation is, even for single-variable equation, the
variable doesn't appear inside any constant.  For example, an equation like
$P \approx E\{P/X\}$ in which $E = A$ and $A \overset{\mathrm{def}}{=}
\tau.A + b.X$ is not supported, because the equation variable $X$
appears inside the definition of constant $A$. On the other side, if
$E = A + X$ with whatever definition of constant $A$ in which there's
no appearence of $X$, such equations are supported. This is not
exactly the same as \emph{pure} equations (no constants anywhere), but
essentially works in the same way. This limitation comes from the fact
that, our congruence definition doesn't consider constants at all.
And to fix the issue, we must prove the Proposition 4.12 in Milner's
book first:
\begin{proposition}
Let $\tilde{E}$ and $\tilde{F}$ contain variables $\tilde{X}$ at
most. Let
$\tilde{A}\overset{\mathrm{def}}{=}\tilde{E}\{\tilde{A}/\tilde{X}\},
\tilde{B}\overset{\mathrm{def}}{=}\tilde{E}\{\tilde{B}/\tilde{X}\}$
and $\tilde{E}\sim\tilde{F}$. Then $\tilde{A}\sim\tilde{B}$.
\end{proposition}
which has one of the longest proof in that book, depends on bisimulation upto
$\sim$ and lemmas on free variables appeared in CCS
processes. This complicated part is planed as the future work.

@@

The congruence (or substitutivity) of bisimilarity relations are
usually defined by a series of substitutivity theorems in most CCS
textbooks (c.f. Milners' and Gorrieri's). The concept of semantic context wasn't explicitly given in
Milner's book.\footnote{Milner says (p.\,84): ``.. congruence
  relation; that is, it is preserved by all algebraic contexts'', but
  he didn't further give a clear definition of algerbaic contexts.}
Later books made it explicitly, for exmaple, in Sangior's book\footnote{Introduction to Bisimulation and
  Coinduction},  a \emph{(semantic) context} is a process expression with a single occurrence of a hold
$[\cdot]$ in it, as a subexpression (c.f. p.99 of Sangiorgi's book).

In van Glabbek's paper\footnote{A characterisation of weak
  bisimulation congruence}), maybe for the first time the concept of
congruence is clearly defined in association with equivalence
relation:
\begin{definition}
An equivalence relation $\approx$ is a congruence for a set of
operators $L$ iff for every $n$-ary operator $f$ in $L$ one has
$g_1\approx h_1 \wedge \cdots \wedge g_n \approx h_n \Rightarrow
f(g_1,\ldots,g_n) \approx f(h_1,\ldots,h_n)$. This is the case iff for
every semantic context $C[\cdot]$ one has $g\approx h \Rightarrow
C[g]\approx C[h]$.
\end{definition}

van Glabbek further gave a way to construct congruence relations from
any existing equivalence relations:
\begin{definition}
Given an equivalence relation $\sim$, its (pre)congruence closure,
notated as $[\sim]$ is defined by
\begin{equation}
g\,[\sim]\, h \mbox{ iff } C[g] \sim C[h] \mbox{ for every semantic
  context } C[\cdot].
\end{equation}
\end{definition}
And it's simple to prove that, for any equivation relation $\sim$,
$[\sim]$ is the coarsest congruence finer than $\sim$.

Here we have basically followed van Glabbek's approach. But the
restriction of single occurrence of the hole seems unnecessary: all
congruence relations in CCS hold also under semantic context with
multiple holes. Further more, a multi-hole semantic context $C$ can be
naturally used for representing an equation (or contraction), e.g. $P
\sim C[P]$.

It remains to find a suitable formal definition of semantic
context. There're multiple ways.\footnote{A literal approach is to
define another algebraic datatype which copies everything from the
existing CCS datatype, plus a dedicated symbol for that ``hole'', then
an inductively defined \texttt{replace} operator can be used for
replacing that hole with any CCS process, returning a normal CCS
datatype.} Here we have chosen to use $\lambda$-expressions (typed
$CCS\rightarrow CCS$) to represent a (multi-hole) semantic
context. The definition is inductive:\footnote{Due to symbol conflicts
  with $\lambda$-calculus, the prefix operator in CCS is represented
  as double-dot in HOL.}
\begin{alltt}
\HOLConst{CONTEXT} (\HOLTokenLambda{}\HOLBoundVar{t}. \HOLBoundVar{t})
\HOLConst{CONTEXT} (\HOLTokenLambda{}\HOLBoundVar{t}. \HOLFreeVar{p})
\HOLConst{CONTEXT} \HOLFreeVar{e} \HOLSymConst{\HOLTokenImp{}} \HOLConst{CONTEXT} (\HOLTokenLambda{}\HOLBoundVar{t}. \HOLFreeVar{a}\HOLSymConst{..}\HOLFreeVar{e} \HOLBoundVar{t})
\HOLConst{CONTEXT} \HOLFreeVar{e\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{CONTEXT} \HOLFreeVar{e\sb{\mathrm{2}}} \HOLSymConst{\HOLTokenImp{}} \HOLConst{CONTEXT} (\HOLTokenLambda{}\HOLBoundVar{t}. \HOLFreeVar{e\sb{\mathrm{1}}} \HOLBoundVar{t} \HOLSymConst{+} \HOLFreeVar{e\sb{\mathrm{2}}} \HOLBoundVar{t})
\HOLConst{CONTEXT} \HOLFreeVar{e\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{CONTEXT} \HOLFreeVar{e\sb{\mathrm{2}}} \HOLSymConst{\HOLTokenImp{}} \HOLConst{CONTEXT} (\HOLTokenLambda{}\HOLBoundVar{t}. \HOLFreeVar{e\sb{\mathrm{1}}} \HOLBoundVar{t} \HOLSymConst{\ensuremath{\parallel}} \HOLFreeVar{e\sb{\mathrm{2}}} \HOLBoundVar{t})
\HOLConst{CONTEXT} \HOLFreeVar{e} \HOLSymConst{\HOLTokenImp{}} \HOLConst{CONTEXT} (\HOLTokenLambda{}\HOLBoundVar{t}. \HOLSymConst{\ensuremath{\nu}} \HOLFreeVar{L} (\HOLFreeVar{e} \HOLBoundVar{t}))
\HOLConst{CONTEXT} \HOLFreeVar{e} \HOLSymConst{\HOLTokenImp{}} \HOLConst{CONTEXT} (\HOLTokenLambda{}\HOLBoundVar{t}. \HOLConst{relab} (\HOLFreeVar{e} \HOLBoundVar{t}) \HOLFreeVar{rf})
\end{alltt}

@@

By induction on the structure of $\lambda$-expressions, we can easily
prove that, a weakly guarded context is also a (normal) context, and the
composition of a (normal) context and a weakly guarded context is
still weakly guarded:
\begin{alltt}
WG_IS_CONTEXT:
\HOLTokenTurnstile{} \HOLConst{WG} \HOLFreeVar{e} \HOLSymConst{\HOLTokenImp{}} \HOLConst{CONTEXT} \HOLFreeVar{e}
CONTEXT_WG_combin:
\HOLTokenTurnstile{} \HOLConst{CONTEXT} \HOLFreeVar{c} \HOLSymConst{\HOLTokenConj{}} \HOLConst{WG} \HOLFreeVar{e} \HOLSymConst{\HOLTokenImp{}} \HOLConst{WG} (\HOLFreeVar{c} \HOLSymConst{\HOLTokenCompose} \HOLFreeVar{e})
\end{alltt}
It's a similar process to define strongly guarded context
(\texttt{SG}), sequential context (\texttt{SEQ}) and their
variants without direct sums (\texttt{GCONTEXT}, \texttt{WGS},
\texttt{GSEQ}). Some lemmas about their relationships have very long
(but trivial) formal proofs due to the combinatorial explosion of
inductions on their structures. For example, one such 
lemma says, for any semantic context $E$ which is both strongly guarded
and sequential (no direct sums) and another sequential context $H$ (no
direct sums), the composition $H \circ E$ is still both strongly
guarded and sequential (no direct sums):
\begin{alltt}
SG_GSEQ_combin:
\HOLTokenTurnstile{} \HOLConst{SG} \HOLFreeVar{E} \HOLSymConst{\HOLTokenConj{}} \HOLConst{GSEQ} \HOLFreeVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{H}. \HOLConst{GSEQ} \HOLBoundVar{H} \HOLSymConst{\HOLTokenImp{}} \HOLConst{SG} (\HOLBoundVar{H} \HOLSymConst{\HOLTokenCompose} \HOLFreeVar{E}) \HOLSymConst{\HOLTokenConj{}} \HOLConst{GSEQ} (\HOLBoundVar{H} \HOLSymConst{\HOLTokenCompose} \HOLFreeVar{E})
\end{alltt}

\subsection{Milner's unique solution of equations theorems}

Once the representation issue of CCS equations is resolved, the actual
proofs of Milner's unique solution of equations theorems is not very
interesting from the view of theorem proving, although it's a 
precise proof engineering work producing quite long proofs.
Since we have chosen to use semantic context to represent
single-variable equations, an equation like $P \sim E\{P/X\}$ now
becomes $P \sim E[P]$, in which $E$ is a (multi-hole) semantic
context, and there's no need to say it ``contains at most the variable
$X$'' any more, as equation variable doesn't appear in $E$ at
all. Under these simplifications, the formal proof of Milner's unique
solution of equations theorem for $\sim$ is a literal mapping for informal
proofs based on bisimulation upto $\sim$, induction and case analysis
of weakly guarded contexts. Below is the formal version of Lemma 4.13
and Proposition 4.14 in Milner's book:
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{WG} \HOLFreeVar{E} \HOLSymConst{\HOLTokenImp{}}
   \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{P} \HOLBoundVar{a} \HOLBoundVar{P\sp{\prime}}.
       \HOLFreeVar{E} \HOLBoundVar{P} \HOLTokenTransBegin\HOLBoundVar{a}\HOLTokenTransEnd \HOLBoundVar{P\sp{\prime}} \HOLSymConst{\HOLTokenImp{}}
       \HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{E\sp{\prime}}. \HOLConst{CONTEXT} \HOLBoundVar{E\sp{\prime}} \HOLSymConst{\HOLTokenConj{}} (\HOLBoundVar{P\sp{\prime}} \HOLSymConst{=} \HOLBoundVar{E\sp{\prime}} \HOLBoundVar{P}) \HOLSymConst{\HOLTokenConj{}} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Q}. \HOLFreeVar{E} \HOLBoundVar{Q} \HOLTokenTransBegin\HOLBoundVar{a}\HOLTokenTransEnd \HOLBoundVar{E\sp{\prime}} \HOLBoundVar{Q}
\HOLTokenTurnstile{} \HOLConst{WG} \HOLFreeVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{P} \HOLBoundVar{Q}. \HOLBoundVar{P} \HOLSymConst{\HOLTokenStrongEQ} \HOLFreeVar{E} \HOLBoundVar{P} \HOLSymConst{\HOLTokenConj{}} \HOLBoundVar{Q} \HOLSymConst{\HOLTokenStrongEQ} \HOLFreeVar{E} \HOLBoundVar{Q} \HOLSymConst{\HOLTokenImp{}} \HOLBoundVar{P} \HOLSymConst{\HOLTokenStrongEQ} \HOLBoundVar{Q}
\end{alltt}

Milner's book only gives the unique solution of equations theorem of
observational congruence ($\approx^c$, Proposition 7.13):
\begin{alltt}
OBS_UNIQUE_SOLUTION:
\HOLTokenTurnstile{} \HOLConst{SG} \HOLFreeVar{E} \HOLSymConst{\HOLTokenConj{}} \HOLConst{SEQ} \HOLFreeVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{P} \HOLBoundVar{Q}. \HOLBoundVar{P} \HOLSymConst{\HOLTokenObsCongr} \HOLFreeVar{E} \HOLBoundVar{P} \HOLSymConst{\HOLTokenConj{}} \HOLBoundVar{Q} \HOLSymConst{\HOLTokenObsCongr} \HOLFreeVar{E} \HOLBoundVar{Q} \HOLSymConst{\HOLTokenImp{}} \HOLBoundVar{P} \HOLSymConst{\HOLTokenObsCongr} \HOLBoundVar{Q}
\end{alltt}
But its original proof is wrong and cannot be fixed even using the
alternative version of bisimulation upto $\approx$ (c.f. The problem
of ``Weak Bisimulation up to''). For this proof, we've made a direct
proof without using any bisimulation upto technique, which essentially
depends on the following relationship between observational congruence
and weak bisimulation (not bisimilarity):
\begin{alltt}
OBS_CONGR_BY_WEAK_BISIM:
\HOLTokenTurnstile{} \HOLConst{WEAK_BISIM} \HOLFreeVar{Wbsm} \HOLSymConst{\HOLTokenImp{}}
   \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{E} \HOLBoundVar{E\sp{\prime}}.
       (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{u}.
            (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{E\sb{\mathrm{1}}}. \HOLBoundVar{E} \HOLTokenTransBegin\HOLBoundVar{u}\HOLTokenTransEnd \HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenImp{}} \HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{E\sb{\mathrm{2}}}. \HOLBoundVar{E\sp{\prime}} \HOLTokenWeakTransBegin\HOLBoundVar{u}\HOLTokenWeakTransEnd \HOLBoundVar{E\sb{\mathrm{2}}} \HOLSymConst{\HOLTokenConj{}} \HOLFreeVar{Wbsm} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLBoundVar{E\sb{\mathrm{2}}}) \HOLSymConst{\HOLTokenConj{}}
            \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{E\sb{\mathrm{2}}}. \HOLBoundVar{E\sp{\prime}} \HOLTokenTransBegin\HOLBoundVar{u}\HOLTokenTransEnd \HOLBoundVar{E\sb{\mathrm{2}}} \HOLSymConst{\HOLTokenImp{}} \HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{E\sb{\mathrm{1}}}. \HOLBoundVar{E} \HOLTokenWeakTransBegin\HOLBoundVar{u}\HOLTokenWeakTransEnd \HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLFreeVar{Wbsm} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLBoundVar{E\sb{\mathrm{2}}}) \HOLSymConst{\HOLTokenImp{}}
       \HOLBoundVar{E} \HOLSymConst{\HOLTokenObsCongr} \HOLBoundVar{E\sp{\prime}}
\end{alltt}

There's actual a third version for weak
bisimilarity not explicitly mentioned in Milner's book, mostly because
it's even more restrictive: beside being strongly guarded and sequential, the
equation variable $X$ must not appear in any direct
sum:\footnote{Here, \texttt{GSEQ} means a sequential context without
  direct sums.}
\begin{alltt}
WEAK_UNIQUE_SOLUTION:
\HOLTokenTurnstile{} \HOLConst{SG} \HOLFreeVar{E} \HOLSymConst{\HOLTokenConj{}} \HOLConst{GSEQ} \HOLFreeVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{P} \HOLBoundVar{Q}. \HOLBoundVar{P} \HOLSymConst{\HOLTokenWeakEQ} \HOLFreeVar{E} \HOLBoundVar{P} \HOLSymConst{\HOLTokenConj{}} \HOLBoundVar{Q} \HOLSymConst{\HOLTokenWeakEQ} \HOLFreeVar{E} \HOLBoundVar{Q} \HOLSymConst{\HOLTokenImp{}} \HOLBoundVar{P} \HOLSymConst{\HOLTokenWeakEQ} \HOLBoundVar{Q}
\end{alltt}

>>>>

In the second half of this paper, we present a formalization of CCS
including the new concepts and theorems proposed in the first half of
this paper. The main purpose is to convince the readers that, there's no flaw in the informal proofs. The
formalization work is based on a formalization of CCS in HOL theorem
prover (HOL4), described in second author's MSc thesis
\cite{Tian:2017wrba} and now lives in the official examples folder of HOL4 source
code\footnote{\url{https://github.com/HOL-Theorem-Prover/HOL}}.
The work is essentially a porting and extension of the early CCS
formalization by Monica Nesi \cite{Nesi} in HOL88.

We will not repeat all definitions and theorems of CCS again by their
formalized versions. Instead, we just focus on several highlights in
this work, i.e.
\begin{enumerate}
\item The formalization of semantics context in $\lambda$-expressions and the theory of
  (pre)congruence for CCS;
\item The definition and uses of trace in the proof of unique solution of
  contractions theorem;
\item The proof of the hard version of ``coarsest congruence
  contained in $\approx$'' theorem.
\end{enumerate}

@@

There're two major limitations in this work. One is the class of CCS
processes, i.e. we're limited with Finitary CCS which has no infinite
sums or parallel compositions. The other limitation is, even for single-variable equation, the
variable doesn't appear inside any constant.

@@

We need to find a suitable formal definition of semantic
context. There're multiple ways. Here we have chosen to use $\lambda$-expressions (typed
$CCS\rightarrow CCS$) to represent a (multi-hole) semantic
context. The definition is inductive:
\begin{alltt}
\HOLConst{CONTEXT} (\HOLTokenLambda{}\HOLBoundVar{t}. \HOLBoundVar{t})
\HOLConst{CONTEXT} (\HOLTokenLambda{}\HOLBoundVar{t}. \HOLFreeVar{p})
\HOLConst{CONTEXT} \HOLFreeVar{e} \HOLSymConst{\HOLTokenImp{}} \HOLConst{CONTEXT} (\HOLTokenLambda{}\HOLBoundVar{t}. \HOLFreeVar{a}\HOLSymConst{..}\HOLFreeVar{e} \HOLBoundVar{t})
\HOLConst{CONTEXT} \HOLFreeVar{e\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{CONTEXT} \HOLFreeVar{e\sb{\mathrm{2}}} \HOLSymConst{\HOLTokenImp{}} \HOLConst{CONTEXT} (\HOLTokenLambda{}\HOLBoundVar{t}. \HOLFreeVar{e\sb{\mathrm{1}}} \HOLBoundVar{t} \HOLSymConst{+} \HOLFreeVar{e\sb{\mathrm{2}}} \HOLBoundVar{t})
\HOLConst{CONTEXT} \HOLFreeVar{e\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{CONTEXT} \HOLFreeVar{e\sb{\mathrm{2}}} \HOLSymConst{\HOLTokenImp{}} \HOLConst{CONTEXT} (\HOLTokenLambda{}\HOLBoundVar{t}. \HOLFreeVar{e\sb{\mathrm{1}}} \HOLBoundVar{t} \HOLSymConst{\ensuremath{\parallel}} \HOLFreeVar{e\sb{\mathrm{2}}} \HOLBoundVar{t})
\HOLConst{CONTEXT} \HOLFreeVar{e} \HOLSymConst{\HOLTokenImp{}} \HOLConst{CONTEXT} (\HOLTokenLambda{}\HOLBoundVar{t}. \HOLSymConst{\ensuremath{\nu}} \HOLFreeVar{L} (\HOLFreeVar{e} \HOLBoundVar{t}))
\HOLConst{CONTEXT} \HOLFreeVar{e} \HOLSymConst{\HOLTokenImp{}} \HOLConst{CONTEXT} (\HOLTokenLambda{}\HOLBoundVar{t}. \HOLConst{relab} (\HOLFreeVar{e} \HOLBoundVar{t}) \HOLFreeVar{rf})
\end{alltt}

@@

It's a similar process to define strongly guarded context
(\texttt{SG}), sequential context (\texttt{SEQ}) and their
variants without direct sums (\texttt{GCONTEXT}, \texttt{WGS},
\texttt{GSEQ}). Some lemmas about their relationships have very long
(but trivial) formal proofs due to the combinatorial explosion of
inductions on their structures. For example, one such 
lemma says, for any semantic context $E$ which is both strongly guarded
and sequential (no direct sums) and another sequential context $H$ (no
direct sums), the composition $H \circ E$ is still both strongly
guarded and sequential (no direct sums):
\begin{alltt}
SG_GSEQ_combin:
\HOLTokenTurnstile{} \HOLConst{SG} \HOLFreeVar{E} \HOLSymConst{\HOLTokenConj{}} \HOLConst{GSEQ} \HOLFreeVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{H}. \HOLConst{GSEQ} \HOLBoundVar{H} \HOLSymConst{\HOLTokenImp{}} \HOLConst{SG} (\HOLBoundVar{H} \HOLSymConst{\HOLTokenCompose} \HOLFreeVar{E}) \HOLSymConst{\HOLTokenConj{}} \HOLConst{GSEQ} (\HOLBoundVar{H} \HOLSymConst{\HOLTokenCompose} \HOLFreeVar{E})
\end{alltt}

\subsection{Milner's ``unique solution of equations'' theorems}

Once the representation issue of CCS equations is resolved, the actual
proofs of Milner's unique solution of equations theorems is not very
interesting from the view of theorem proving, although it's a 
precise proof engineering work producing quite long proofs.
Since we have chosen to use semantic context to represent
single-variable equations, an equation like $P \sim E\{P/X\}$ now
becomes $P \sim E[P]$, in which $E$ is a (multi-hole) semantic
context, and there's no need to say it ``contains at most the variable
$X$'' any more, as equation variable doesn't appear in $E$ at
all. Under these simplifications, the formal proof of Milner's unique
solution of equations theorem for $\sim$ is a literal mapping for informal
proofs based on bisimulation upto $\sim$, induction and case analysis
of weakly guarded contexts. Below is the formal version of Lemma 4.13
and Proposition 4.14 in Milner's book:

\begin{alltt}
STRONG_UNIQUE_SOLUTION:
\HOLTokenTurnstile{} \HOLConst{WG} \HOLFreeVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{P} \HOLBoundVar{Q}. \HOLBoundVar{P} \HOLSymConst{\HOLTokenStrongEQ} \HOLFreeVar{E} \HOLBoundVar{P} \HOLSymConst{\HOLTokenConj{}} \HOLBoundVar{Q} \HOLSymConst{\HOLTokenStrongEQ} \HOLFreeVar{E} \HOLBoundVar{Q} \HOLSymConst{\HOLTokenImp{}} \HOLBoundVar{P} \HOLSymConst{\HOLTokenStrongEQ} \HOLBoundVar{Q}
\end{alltt}

\begin{alltt}
WEAK_UNIQUE_SOLUTION:
\HOLTokenTurnstile{} \HOLConst{SG} \HOLFreeVar{E} \HOLSymConst{\HOLTokenConj{}} \HOLConst{GSEQ} \HOLFreeVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{P} \HOLBoundVar{Q}. \HOLBoundVar{P} \HOLSymConst{\HOLTokenWeakEQ} \HOLFreeVar{E} \HOLBoundVar{P} \HOLSymConst{\HOLTokenConj{}} \HOLBoundVar{Q} \HOLSymConst{\HOLTokenWeakEQ} \HOLFreeVar{E} \HOLBoundVar{Q} \HOLSymConst{\HOLTokenImp{}} \HOLBoundVar{P} \HOLSymConst{\HOLTokenWeakEQ} \HOLBoundVar{Q}
\end{alltt}

\begin{alltt}
OBS_UNIQUE_SOLUTION:
\HOLTokenTurnstile{} \HOLConst{SG} \HOLFreeVar{E} \HOLSymConst{\HOLTokenConj{}} \HOLConst{SEQ} \HOLFreeVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{P} \HOLBoundVar{Q}. \HOLBoundVar{P} \HOLSymConst{\HOLTokenObsCongr} \HOLFreeVar{E} \HOLBoundVar{P} \HOLSymConst{\HOLTokenConj{}} \HOLBoundVar{Q} \HOLSymConst{\HOLTokenObsCongr} \HOLFreeVar{E} \HOLBoundVar{Q} \HOLSymConst{\HOLTokenImp{}} \HOLBoundVar{P} \HOLSymConst{\HOLTokenObsCongr} \HOLBoundVar{Q}
\end{alltt}
